\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-12-30 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

{\nib \blue{TBD.}}

\noindent
{\bf R4.3. Comments by reviewer}:\\
{\em Regarding the presentation, the reviewer suggests making the paper more self-contained. Some discussion in this paper highly relies on the authors' previous work, which can be improved by either adding a summarization or trimming out the materials.

Following the same vein, there are some unnecessary information and technical points provided in this paper that may be distracting.
For example, on page 3, why would the authors mention the Kuramoto oscillator model?
On page 4, is the detailed discussion regarding Wasserstein-2 metric necessary? Similarly, on page 5, the discussion about the natural energy functional and how the gradient drift case does not apply also seems too detailed.}

{\nib \blue{We have followed the reviewer's suggestion to improve the presentation. \ul{Specifically, in the revised manuscript's Sec. III-A, last two paragraphs summarize the technical relation of this paper w.r.t. the authors' previous work. In particular, we have added three more sentences with a new reference and broken what was a single paragraph in the previous version into two paragraphs.}

@@

We appreciate the reviewer's point that the sentence containing equation (6) and the mention of Kuramoto model can be removed without affecting the rest of the development. However, we prefer to keep it since it adds didactic value, and many engineers may feel comfortable by the explicit mention that (1) is simply a rewriting of (6). Furthermore, an astute reader, encountering that sentence will realize that the scope of the technical contribution of this paper is broad because the algorithmic framework developed here will be applicable to several other non-power system applications (mentioned in the two cited references in that sentence) where the second order Kuramoto model is used. Since this is just a matter of one sentence, we will prefer to keep it. We hope you approve.

\ul{On the other hand, we believe the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}. Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory and simulations. \ul{The key contribution of this paper, from the authors' viewpoint, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature}. We believe this clarification of what really is the technical crux of the matter--versus what are peripheral issues--should be illuminating to not just power systems researchers, but also to applied mathematicians underscoring how the standard theory and algorithms for measure-valued gradient flows need to be adapted in engineering applications.}}

### 2021-12-30 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
TPS  rebuttal round 1

The algorithmic appeal of the proximal recursion \eqref{ProxRecursionInfiniteDim} is that it opens up the possibility to compute the solution of the PDE initial value problem via recursive convex minimization. A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime. Notice that even though the drift in \eqref{ItoSDEvectorlevel} is \emph{not} a gradient vector field, the algorithm in \cite[Sec. V.B]{caluya2019TAC} constructed a pair $\left(\Phi,{\rm{dist}}\right)$ such that \eqref{ProxRecursionInfiniteDim} provably approximates the transient solution of the corresponding kinetic Fokker-Planck PDE with guarantee \eqref{ConvergenceGuarantee}. However, that algorithm cannot be applied to \eqref{KineticFPK} as is. The reasons are explained next.

@@

\bibitem{villani2003topics}
C. Villani, \emph{Topics in optimal transportation}, No. 58, American Mathematical Society, 2003.

@@

>>>>

The algorithmic appeal of the proximal recursion \eqref{ProxRecursionInfiniteDim} is that it opens up the possibility to compute the solution of the PDE initial value problem via recursive convex minimization. A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime. {\blue{The main idea in these references was to \emph{co-evolve} the time-varying state samples (via numerical SDE solver) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples. The resulting computation is an \emph{online} propagation of the joint PDFs as opposed to the computation in Monte Carlo or density estimation methods \cite{silverman1998density}. The latter methods only propagate the state samples and then approximate the joint PDFs as \emph{offline} post-processing.}}

Notice that even though the drift in \eqref{ItoSDEvectorlevel} is \emph{not} a gradient vector field, the algorithm in \cite[Sec. V.B]{caluya2019TAC} constructed a pair $\left(\Phi,{\rm{dist}}\right)$ such that \eqref{ProxRecursionInfiniteDim} provably approximates the transient solution of the corresponding kinetic Fokker-Planck PDE with guarantee \eqref{ConvergenceGuarantee}. However, that algorithm cannot be applied to \eqref{KineticFPK} as is. The reasons are explained next.

@@

\bibitem{villani2003topics}
C. Villani, \emph{Topics in Optimal Transportation}, No. 58, American Mathematical Society, 2003.

@@

% === density estimation ====

\bibitem{silverman1998density}
{\blue{B.W.~Silverman, \emph{Density Estimation for Statistics and Data Analysis}, 1st ed., Chapman \& Hall/CRC, 1998.}}
